---
output:
  html_document: default 
  pdf_document: default
---

---
title:  'Twitter Retweetability Analysis'
author:
- Julio Rengifo
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

# Problem
The aim of this project is to understand the relationship between retweetability of one tweet (whether the tweet is retweeted and, if it does, how many times it is retweeted) and some features of the tweet. 


# Data
```{r message=FALSE, warning=FALSE}
# import data
install.packages("readr")
library(readr)
LLBean_retweet <- read_csv("LLBean_retweet.csv")

# wrangle data
install.packages(c("dplyr", "stringr","stringi"))
library(dplyr)
LLBean1<-LLBean_retweet%>%filter(language == c('en'))%>%
        mutate(hashtags = gsub("\\[|\\]", "", hashtags), 
               urls=gsub("\\[|\\]", "", urls))

library(stringr)
library(stringi)
LLBean2<-LLBean1%>% mutate(
tweet_length=str_length(tweet),
url_ind=ifelse(str_length(urls)==0, 0, 1),                                             hashtags_count=ifelse(str_length(hashtags)==0,0,stri_count_fixed(hashtags, ",") + 1),
retweet_ind=as.numeric(retweets_count>0))%>%select(-language)

LLBean3<-LLBean2%>%filter(retweet_ind==1) 

head(LLBean_retweet)
head(LLBean1)
head(LLBean2)
head(LLBean3)
```

# Analysis
## Logistic regression model
```{r message=FALSE, warning=FALSE}
install.packages("caret")
library(caret)
install.packages("car")
library(car)

# set up lgr formula
LLBean2_lgr <- glm(retweet_ind~tweet_length+url_ind+hashtags_count+video,
                data = LLBean2,
                family = binomial(link = logit))
LLBean2_lgr
# address the concern of multicollinearity 

vif(LLBean2_lgr) #the vif() output indicates that there is no significant multicollinearity.

# display the importance of different factors
varImp(LLBean2_lgr, scale = TRUE)

```

**Based on my final model result it is apparent that video is the  the most significant predictor of retweet_ind (all else equal), with every one-unit increase in in video the log probability of retweet_ind increases by a factor of 1.36**

```{r message=FALSE, warning=FALSE}
confint(LLBean2_lgr, "video", level = 0.95)
```


**Based on this model and the assumption of causality relationship between retweet_ind and other predictors in the logit regression, we predict with 95% confidence that for every one unit increase in video the odds of a retweet_ind outcome will increase by a possible value in the range between 133% to 139%.**

```{r message=FALSE, warning=FALSE}
install.packages("DescTools")
library(DescTools)
Cstat(LLBean2_lgr)
```

**Based on the Cstat of .673977 the performance of the model is average**

## Least-square regression model
```{r message=FALSE, warning=FALSE}
install.packages("gvlma")
install.packages("car")
library(gvlma)
library(car)

LLBean3_lsr <- lm(formula = retweets_count~tweet_length+url_ind+hashtags_count+video, 
                  data = LLBean3)
LLBean3_lsr

gvlma(LLBean3_lsr, alphalevel = 0.001)

plot(LLBean3_lsr, 4)

#Calculate Cook's distance and identify records with a Cook's distance high enough to warrant removing form the dataset
cd <- cooks.distance(LLBean3_lsr)
high_cd <- which(cd > 4 * mean(cd))

# remove records with high Cook's distance
LLBean3_updated <- LLBean3[-high_cd,]

LLBean3_lsr_new <- lm(formula = retweets_count~tweet_length+url_ind+hashtags_count+video, 
                  data = LLBean3_updated)
LLBean3_lsr_new

gvlma(LLBean3_lsr_new, alphalevel = 0.001)

plot(LLBean3_lsr_new, 4)

#check for multicollinearity 
vif(LLBean3_lsr_new)

# display the importance of different factors
varImp(LLBean3_lsr_new, scale = TRUE)
```

**global stat, skewness, kurtosis, and heteroscedasticity are not satisfied using an OLS regression model on this dataset**


**Based on my final model result it is apparent that video is the  the most significant predictor of retweet_ind (all else equal), with every one-unit increase in in video the log probability of retweet_ind increases by a factor of 1.716878**

```{r message=FALSE, warning=FALSE}
confint(LLBean3_lsr_new, "video", level = 0.95)
```

**Based on this model and the assumption of causality relationship between retweet_ind and other predictors in the least squares regression, we predict with 95% confidence that for every one unit increase in video the odds of a retweet_ind outcome will increase by a possible value in the range between 154% to 189%.**

```{r message=FALSE, warning=FALSE}
summary(LLBean3_lsr_new)$adj.r.squared
```


**I do not think that the model is a good model, it may be inappropriate for the dataset being observed. the adjusted R-squared tells me that the model explains an insignificant amount of variation in the response variable.**

